---
title: "Prac4"
format: html
editor: visual
execute:
  echo: true
  warning: false
  message: false
---

## Prac 4

### TidyVerse (Question 1-6)

```{r}
library(tidyverse)
library(nycflights13)
library(dplyr)

#Question 1
flights

#Question 2
ans = flights %>% 
  filter(month == 1) %>% # find specific row
  group_by(carrier) %>% 
  summarise(carrier_dist_vec_mean = mean(distance), 
            carrier_dist_vec_sd = sd(distance)) %>% 
  arrange(carrier_dist_vec_mean) # order by = arrange
ans

#Question 3
# You cannot have the sd of 1 value and in month 1 OO only has 1 occurance
flights %>% 
  filter(month == 1) %>% 
  filter(carrier == "OO")

# The 0s arise from all the sd of YV, F9, AS, HA all have multiple identical data points
flights %>% 
  filter(month == 1, carrier == "YV") %>% 
  select(distance)

flights %>% 
  filter(month == 1, carrier == "F9") %>% 
  select(distance)

flights %>% 
  filter(month == 1, carrier == "AS") %>% 
  select(distance)

flights %>% 
  filter(month == 1, carrier == "HA") %>% 
  select(distance)

# Question 4
flights %>% 
  group_by(carrier, month) %>% 
  summarise(avg_delay = mean(dep_delay, na.rm = TRUE)) %>% 
  pivot_wider(
    names_from = carrier,
    values_from = avg_delay
  )

# Question 5
count = flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay)) %>% 
  summarise(nTot = n())

prop = flights %>% 
  filter(dep_delay > 0, arr_delay <= 0, !is.na(dep_delay), !is.na(arr_delay)) %>% 
  summarise(nSat = n())

ans = prop$nSat / count$nTot
ans

# One liner
ans <- flights %>%
  filter(!is.na(dep_delay), !is.na(arr_delay)) %>%
  summarise(proportion = mean(dep_delay > 0 & arr_delay <= 0, na.rm = TRUE))
ans
# This one liner create a true false vector and then 
# just sum up the trues and divide by total

#Question 6
Q1 = flights %>% 
  distinct(origin, dest, carrier) %>% 
  count(origin, dest, name = "numCarrier") %>% 
  filter(numCarrier > 1)
Q1

avg = flights %>% 
  group_by(origin, dest, carrier) %>% 
  summarise(avg_arr_delay = mean(arr_delay, na.rm = TRUE))

Q2 = avg %>% 
  group_by(origin, dest) %>% 
  filter(n() > 1) %>%
  left_join(airlines, by = "carrier")
Q2

Q3 = Q2 %>%
  group_by(origin, dest) %>% 
  summarise(worst_arrival = max(avg_arr_delay),
            worst_airline = name[which.max(avg_arr_delay)],
            best_arrival = min(avg_arr_delay), 
            best_airline = name[which.min(avg_arr_delay)]) %>% 
  ungroup()
Q3
#group_vars(Q3)

Q4 = Q3 %>%
  mutate(diff = worst_arrival - best_arrival) %>% 
  arrange(desc(diff)) %>% 
  slice(1)
Q4

Q5 = flights %>% 
  left_join(airlines, by = "carrier") %>% 
  filter(origin == "JFK", dest == "ATL", name == "ExpressJet Airlines Inc.") 
Q5

```
The explaination for 6.5 is basically because it has the highest absolute difference. ExpressJet only ever does this route once (as shown above) and it had a huge delay and thus there is no way of pulling down the delay time as it is a single value. Therefore the law of averages cannot pull down the large delay. So when comparing it to the flight on this route with the best mean itll create a large difference.

### Question 7

```{r}
dat = structure(list(id = c("id_1", "id_2", "id_3", "id_4", "id_5", 
                      "id_6", "id_7", "id_8", "id_9", "id_10", "id_11", "id_12", "id_13", 
                      "id_14", "id_15", "id_16", "id_17", "id_18", "id_19", "id_20", 
                      "id_21", "id_22", "id_23", "id_24", "id_25", "id_26", "id_27", 
                      "id_28", "id_29", "id_30", "id_31", "id_32", "id_33", "id_34", 
                      "id_35", "id_36", "id_37", "id_38", "id_39", "id_40", "id_41", 
                      "id_42", "id_43", "id_44", "id_45", "id_46", "id_47", "id_48", 
                      "id_49", "id_50"), 
               age = c(50L, 34L, 70L, 33L, 22L, 61L, 69L, 
                       73L, 62L, 56L, 71L, 33L, 73L, 44L, 45L, 46L, 24L, 70L, 46L, 76L, 
                       47L, 76L, 28L, 48L, 54L, 27L, 45L, 26L, 61L, 28L, 38L, 55L, 33L, 
                       36L, 62L, 58L, 72L, 31L, 34L, 51L, 61L, 64L, 26L, 28L, 60L, 29L, 
                       42L, 46L, 79L, 72L), 
               gender = c("male", "male", "male", "female", 
                          "female", "male", "female", "male", "male", "female", "female", 
                          "male", "male", "female", "male", "male", "male", "male", "female", 
                          "male", "male", "male", "male", "female", "femal", "male", "female", 
                          "female", "female", "female", "male", "female", "female", "female", 
                          "male", "male", "female", "male", "female", "female", "male", 
                          "female", "female", "male", "male", "female", "male", "male", 
                          "male", "female"), 
               height = c(174.4, 197.7, 174.1, 194.5, NA, 
                          180.4, 170.5, 157.4, 196.8, 165.1, 153, 197.4, 186, 157.1, 177.5, 
                          197.7, 179.3, 170.2, 182.4, NA, 165.4, 161, 168.5, 199.2, 157.7, 
                          154.6, 157.1, 184.5, 181, 194.6, 183.6, 186.9, 176.1, 183, 191.1, 
                          189.3, 199, 172, 165.6, 170.5, 150.5, 159.2, 192.1, 161.6, 162, 
                          153.8, 162.3, 186.6, 192.4, 174.9), 
               weight = c(69.4, 62.3, 55.6, 
                          69.5, 78.6, 60.8, 72.2, 60.9, 75.1, 67.7, 82.5, 68.7, 67.8, 76.7, 
                          87, 61.1, 70.6, 63.3, 81.5, 59.2, 93.2, 87.3, 83.4, 80.9, 68.6, 
                          76.5, 93.7, 79.1, 92, 65.6, 85.4, 63.3, 79.7, 74.1, 63.3, 78.2, 
                          95.7, 95.1, 63.7, 66.1, 99.3, 81, 96.9, 73.3, 70.3, 83, 57.6, 
                          78.6, 61.9, 98.1), 
               blood_type = c("O", "A", "O", "O", "B", "AB", 
                              "O", "O", "O", "AB", "A", "O", "O", "O", "B", "A", "B", "AB", 
                              "O", "AB", "A", "AB", "O", "B", "A", "A", "B", "AB", "A", "B", 
                              "B", "A", "O", "O", "O", "B", "O", "A", "A", "B", "A", "O", "AB", 
                              "A", "A", "O", "O", "B", "A", "O"), 
               disease_status = c("diseased", 
                                  "healthy", "healthy", "healthy", "healthy", "healthy", "diseased", 
                                  "healthy", "diseased", "Healthy", "diseased", "healthy", "diseased", 
                                  "healthy", "diseased", "healthy", "healthy", "healthy", "healthy", 
                                  "healthy", "healthy", "diseased", "healthy", "diseased", "healthy", 
                                  "healthy", "healthy", "healthy", "diseased", "diseased", "healthy", 
                                  "healthy", "healthy", "diseased", "diseased", "diseased", "healthy", 
                                  "diseased", "healthy", "healthy", "healthy", "healthy", "healthy", 
                                  "diseased", "diseased", "diseased", "healthy", "healthy", "diseased", 
                                  "diseased"), 
               cholesterol = c(228, 223, 213, 198, 166, 151, 195, 
                               199, 189, 196, 221, 156, 185, 230, 234, 174, 185, 236, 235, 180, 
                               165, 220, 160, 153, 250, 153, 184, 242, 212, 179, 224, 233, 181, 
                               199, 220, 214, 214, 248, 191, 162, 203, 173, 199, 187, 248, 189, 
                               173, 212, 164, 247), 
               glucose = c(96, 78, 101, 119, 103, 91, 86, 
                           NA, 77, 80, 115, 85, 88, 109, NA, 71, 90, 94, 91, 87, 113, 93, 
                           97, 118, 109, 80, 85, 119, 99, 108, 89, 108, 97, 116, 79, 84, 
                           75, 81, 119, NA, 106, 109, 75, 82, 84, 75, 76, 120, 119, 77), 
               smoker = c("yes", "yes", "yes", "yes", "no", "yes", "no", 
                          "yes", "no", "no", "no", "no", "no", "yes", "no", "yes", 
                          "yes", "yes", "yes", "yes", "yes", "yes", "yes", "yes", "no", 
                          "no", "yes", "yes", "yes", "no", "no", "yes", "no", "yes", 
                          "no", "yes", "no", "yes", "yes", "yes", "no", "no", "yes", 
                          "no", "no", "no", "no", "no", "no", "yes"), 
               exercise = c("occasional", 
                            "regular", "occasional", "regular", "none", "occasional", 
                            "regular", "none", "occasional", "none", "occasional", "none", 
                            "none", "regular", "occasional", "none", "regular", "regular", 
                            "none", "occasional", "none", "occasional", "occasional", 
                            "occasional", "regular", "occasional", "regular", "regular", 
                            "regular", "occasional", "occasional", "none", "none", "regular", 
                            "occasional", "occasional", "none", "none", "none", "none", 
                            "occasional", "regular", "regular", "none", "regular", "occasional", 
                            "occasional", "none", "occasional", "regular"), 
               income = c(84820L, 
                          81547L, 22588L, 72490L, 74533L, 25338L, 41469L, 57315L, 63629L, 
                          88662L, 62615L, 56261L, 58499L, 82232L, 77584L, 77275L, 38468L, 
                          54510L, 91326L, 78611L, 31402L, 29586L, 21441L, 58269L, 84173L, 
                          88295L, 37940L, 43750L, 69750L, 92356L, 82518L, 91455L, 68866L, 
                          51178L, 68275L, 27689L, 35418L, 81318L, 62405L, 86851L, 25654L, 
                          47553L, 74474L, 51409L, 22607L, 55360L, 96351L, 21516L, 41927L, 
                          55810L), 
               education = c("master", "bachelor", "PhD", "master", 
                             "bachelor", "highschool", "PhD", "highschool", "PhD", "PhD", 
                             "bachelor", "highschool", "master", "bachelor", "PhD", "PhD", 
                             "PhD", "bachelor", "master", "highschool", "PhD", "highschool", 
                             "bachelor", "master", "highschool", "highschool", "master", 
                             "master", "bachelor", "PhD", "highschool", "PhD", "master", 
                             "master", "master", "PhD", "highschool", "master", "master", 
                             "highschool", "bachelor", "highschool", "bachelor", "PhD", 
                             "bachelor", "highschool", "master", "highschool", "bachelor", 
                             "bachelor"), 
               region = c("North", "South", "North", "West", 
                         "North", "West", "South", "South", "West", "South", "West", 
                         "South", "West", "East", "North", "West", "North", "North", 
                         "West", "North", "East", "West", "South", "North", "North", 
                         "East", "East", "North", "North", "West", "South", "West", 
                         "West", "East", "West", "North", "West", "North", "East", 
                         "North", "West", "South", "South", "East", "North", "West", 
                         "West", "East", "North", "East"), 
               marital_status = c("divorced", 
                                  "single", "divorced", "divorced", "divorced", "divorced", 
                                  "divorced", "married", "divorced", "married", "divorced", 
                                  "widowed", "married", "single", "widowed", "widowed", "single", 
                                  "divorced", "widowed", "widowed", "single", "married", "single", 
                                  "married", "widowed", "married", "single", "single", "widowed", 
                                  "married", "widowed", "divorced", "single", "married", "single", 
                                  "widowed", "widowed", "married", "widowed", "divorced", "married", 
                                  "married", "divorced", "single", "married", "widowed", "divorced", 
                                  "divorced", "single", "divorced")), 
          row.names = c(NA, -50L
                                                                                                                                                                                                                                                  
                        ), 
          class = c("tbl_df", "tbl", "data.frame"))
#N/As
#Nulls = dat %>% filter(if_any(everything(), is.na))
Nulls = dat %>% 
  mutate(rowNum = row_number()) %>% 
  filter(if_any(everything(), is.na))
# First check to see where inconsistencies are
# dat %>% count(gender) check individual
# Check all in freq table:
irreg = dat %>%
  select(gender, disease_status, education, smoker, exercise, region, marital_status) %>% 
  pivot_longer(everything(), names_to = "vars", values_to = "vals") %>% 
  count(vars, vals)
  # Now I know that disease_status and gender have mispellings
genderErr = dat %>% 
  mutate(rowNum = row_number()) %>% 
  filter(gender == "femal")

diseaseErr = dat %>% 
  mutate(rowNum = row_number()) %>% 
  filter(disease_status == "Healthy")

Q7 = bind_rows(Nulls, genderErr, diseaseErr)
Q7
# if (!requireNamespace("remotes", quietly = TRUE)) {
#   install.packages("remotes")
# }

# remotes::install_github("SATVILab/UtilsDataRSV")
# UtilsDataRSV::view_cols(dat)
```

\

The `echo: false` option disables the printing of code (only output is displayed).
